# Bayesian analysis of cars and mechanics


We wish to know the probability that if the mechanic *says* a car is
“OK,” it *really* is faulty. Phrased differently, what is the
probability of a car being faulty if the mechanic said it was OK?

```{r}
n_trials <- 10000  # number of cars

# Counters for number of approved, number of approved and faulty
approved <- 0
approved_and_faulty <- 0

for (i in 1:n_trials) {

    # Decide whether the car is faulty or working, with a probability of
    # 0.3 and 0.7 respectively
    car <- sample(c('faulty', 'working'), size=1, prob=c(0.3, 0.7))

    if (car == 'faulty') {
        # What the mechanic says of a faulty car
        mechanic_says <- sample(c('approved', 'not approved'),
                                size=1,
                                prob=c(0.1, 0.9))
    } else {
        # What the mechanic says of a working car
        mechanic_says <- sample(c('approved', 'not approved'),
                                size=1,
                                prob=c(0.7, 0.3))
    }

    if (mechanic_says == 'approved') {
        approved <-  approved + 1
        if (car == 'faulty') {
            approved_and_faulty <- approved_and_faulty + 1
        }
    }
}

k <- approved_and_faulty / approved

message('Proportion of faulty cars of cars approved: ', round(k, 2))
```

The answer looks to be somewhere between 5 and 6%. The code clearly
follows the description step by step, but it is also quite slow. If we
can improve the code, we may be able to do our simulation with more
cars, and get a more accurate answer.

Let’s use arrays to store the states of all cars in the lot
simultaneously:

```{r}
# Number of cars; we made this number larger by a factor of 100
n_trials <- 1000000

# Generate an array with as many entries as there are cars, each
# being either 'working' or 'faulty'.
# We are taking a sample _with_ replacement.
cars <- sample(c('working', 'faulty'),
               size=n_trials,
               replace=TRUE,
               prob=c(0.7, 0.3))

# Count how many cars are working
n_working <- sum(cars == 'working')

# All the rest are faulty
n_faulty <- n_trials - n_working

# Create a new vector in which to store what a mechanic says
# about the car: 'approved' or 'not approved'. We use
# "character" to tell R these are strings.
mechanic_says <- character(n_trials)

# We start with the working cars; what does the mechanic say about them?
# Generate 'approved' or 'not approved' labels with the given probabilities.
mechanic_says[cars == 'working'] <- sample(
    c('approved', 'not approved'),
    size=n_working,
    replace=TRUE,
    prob=c(0.8, 0.2)
)

# Similarly, for each faulty car, generate 'approved'/'not approved'
# labels with the given probabilities.
mechanic_says[cars == 'faulty'] <- sample(
    c('approved', 'not approved'),
    size=n_faulty,
    replace=TRUE,
    prob=c(0.1, 0.9)
)

# Identify all cars that were approved
# This produces a binary mask, an array that looks like:
# [True, False, False, True, ... ]
approved <- (mechanic_says == 'approved')

# Identify cars that are faulty AND were approved
faulty_but_approved <- (cars == 'faulty') & approved

# Count the number of cars that are faulty but approved, as well as
# the total number of cars that were approved
n_faulty_but_approved <- sum(faulty_but_approved)
n_approved <- sum(approved)

# Calculate the ratio, which is the answer we seek
k <- n_faulty_but_approved / n_approved

message('Proportion of faulty cars of cars approved: ', round(k, 2))
```

The code now runs much faster, and with a larger number of cars we see
that the answer is closer to a 5% chance of a car being broken after it
has been approved by a mechanic.
