# Bootstrap test of birthweight difference


We run bootstrap resampling from pooled drug and control birthweights,
to establish the null-world distribution of the mean difference in
weights. Then we compare the observed difference in means to the
null-world distribution to ask if the actual value is plausible in the
null world.

Proceed with the simulation:

```{r}
# Get treatment and control values from data file.
birth_df <- read.csv('data/birthweights.csv')
# Birthweidhts for Drug A participants.
treat_df <- subset(birth_df, birth_df$Treatment == 'Drug A')
# Birthweigts as vector.
treat <- treat_df$Birthweight
# Control birthweights.
control_df <- subset(birth_df, birth_df$Treatment == 'Control')
# Birthweigts as vector.
control <- control_df$Birthweight

# Actual difference.
actual_diff <- mean(treat) - mean(control)
# Show the actual difference.
actual_diff
```

Proceed with the simulation:

```{r}
# Concatenate treatment and control vectors.
both <- c(treat, control)

# Number of trials.
n_trials <- 10000

# Make vector to store results for each trial.
results <- numeric(n_trials)

# Do 10000 simulations
for (i in 1:n_trials) {
    # Take a resample of 15 from all birth weights.
    fake_treat <- sample(both, size=15, replace=TRUE)
    # Take a second, similar resample.
    fake_control <- sample(both, size=15, replace=TRUE)
    # Find the means of the two resamples.
    mt <- mean(fake_treat)
    mc <- mean(fake_control)
    # Find the difference between the means of the two resamples.
    diff <- mt - mc
    # Keep score of the result.
    results[i] <- diff
    # End the simulation experiment, go back and repeat
}
# Produce a histogram of the resample differences
hist(results, breaks=25,
    main='Null-world distribution of treatment/control difference',
    xlab='Null-world mean treatment - mean control')
# How often did resample differences exceed the observed difference of
# .82?
k <- sum(results >= actual_diff)
kk <- k / n_trials
message('Proportion null-world differences >= actual difference: ', kk)
```
