# Building strings for labels


As the examples in this book proceed, we will use more code techniques
to write the examples in a more concise and efficient way.

One task that we often have, is to build up helpful strings to use as
labels on plots, or to print out as messages. These strings will often
mix numbers and text. For example, we may want to print out a helpful
message such as: `Simulation using 10000 trials`, where the number 10000
in the message comes from some variable, such as `n_trials`. Let’s set
that variable now:

```{r}
n_trials <- 10000
```

Building the useful string above would involve taking the string
`'Simulation using '`, then appending a string to represent the number
10,000 — as in:

```{r}
# Convert the number (integer) to a string (character vector) representing the
# number.
as.character(n_trials)
```

Finally we need to append another string to the result — `' trials.'`.
So our task will be to *concatenate* (stick together) these three
strings.

You can concatenate strings in R with the `paste` function. We pass
`paste` the strings and values we want it to combine, as arguments. It
converts each argument to a string, and concatenates the resulting
strings, separating them with a space (by default).

For example, to make a new string that concatenates the strings
`'resampling'`, `'is'` and `'better'`, separated by spaces, we could
use:

```{r}
paste('resampling', 'is', 'better')
```

If we wanted to insert a number as a string into this sequence, we can
send the *number* as an argument. `paste` converts the number to a
string before concatenation.

```{r}
paste('resampling', 'is', 100, 'times', 'better')
```

To create the string we started this section with, we could write:

```{r}
paste('Simulation using', n_trials, 'trials.')
```

We will start using `paste` from here, to compile messages to print, or
labels for plots.


