# Testing association of compliance and cholesterol values


We have 164 rows, each representing a person taking a medication. For
each row (each person) we have a pair of measurements. The first is
their percent compliance in taking their prescribed medication. The
second is the decrease of their blood cholesterol. We want to see
whether the observed association of these values is plausible on the
null hypothesis — that there is no association.

```{r}
# Data as vectors.
cholost_df <- read.csv('data/cholost.csv')
x <- cholost_df$percent_compliance
y <- cholost_df$cholesterol_decrease

# Step 1 above
actual_prod <- x * y
# Note: actual_sum <- 439,141
actual_sum <- sum(actual_prod)
# Show the result.
actual_sum
```

```{r}
# Set the number of trials
n_trials <- 10000

# An empty array to store the trial results.
results <- numeric(n_trials)

# Do 10,000 experiments (step 4)
for (i in 1:n_trials) {
    # Step 2 above.
    y_random <- sample(y)
    # Step 3 above.
    fake_prod <- x * y_random
    fake_sum <- sum(fake_prod)
    # Step 3 above
    results[i] <- fake_sum
    # Step 4 above
}

# Step 5 above
k <- sum(results >= actual_sum)
kk <- k / n_trials

message('Proportion product sums >= observed: ', kk)
```
