# Permutation test of public and private liquor prices


This notebook asks the question whether the difference in the means of
private and government-specified prices of a particular whiskey could
plausibly have come about as a result of random sampling. It uses a
permutation method, where we use a shuffled version of all 42 values to
comprise our null-world samples.

```{r}
# Load the data from a data file.
prices_df <- read.csv('data/liquor_prices.csv')
# Show this first six rows.
head(prices_df)
```

Take all prices from the loaded data file, and convert into a vectors
for each category.

```{r}
# Rows for private prices.
priv_df <- subset(prices_df, prices_df['state_type'] == 'private')
# Convert corresponding prices to vector.
priv <- priv_df$price
# Show the result.
priv
```

```{r}
# Rows for government prices.
govt_df <- subset(prices_df, prices_df['state_type'] == 'government')
# Convert corresponding prices to vector.
govt = govt_df$price
# Show the result.
govt
```

Calculate actual difference:

```{r}
actual_diff <- mean(priv) - mean(govt)
actual_diff
```

Concatenate the private and government values into one vector:

```{r}
# Join the two vectors of data into one vector.
both <- c(priv, govt)
both
```

Do simulation:

```{r}
n_trials <- 10000

# Fake differences for each trial.
results <- numeric(n_trials)

# Repeat 10000 simulation trials
for (i in 1:10000) {

    # Shuffle 42 values to a random order.
    shuffled = sample(both)

    # Take first 26 shuffled values as fake private group
    fake_priv = shuffled[1:26]

    # Remaining values (from position 27 to end, 16 values)
    # form the fake government group.
    fake_govt = shuffled[27:42]

    # Find the mean of the "private" group.
    p <- mean(fake_priv)

    # Mean of the "govt." group
    g <- mean(fake_govt)

    # Difference in the means
    diff <- p - g

    # Keep score of the trials
    results[i] <- diff
}

# Graph of simulation results to compare with the observed result.
fig_title <- paste('Average price difference (Actual difference = ',
                   round(actual_diff * 100),
                   'cents')
hist(results, main=fig_title, xlab='Difference in average prices (cents)')

# Number of trials where fake difference >= actual.
k = sum(results >= actual_diff)
kk = k / n_trials

message('Proportion fake differences <= actual_difference: ', kk)
```
